#import all necessary libraries
```{r libraries}
library(readr)
library(dplyr)
library(ggplot2)
library(plotly)

```


#this block is used to read in the data
```{r data}
gym_members_exercise_tracking <- read_csv("gym_members_exercise_tracking.csv")
```

#this block is used to make all necessary conversions for the data
```{r pre-processing}
#creates a new data frame that all further manipulations will be perfomed on
gym = gym_members_exercise_tracking

#converts the weight into pounds from kilograms 
#renames it to "Weight (lbs)"
gym = gym %>%
  mutate(`Weight (kg)` = round(`Weight (kg)` * 2.20462, 1))
gym = gym %>%
  rename(`Weight (lbs)` = `Weight (kg)`)

#converts the height into inches from meters 
#renames it to "Height (in)"
gym = gym %>% 
  mutate(`Height (m)` = round(`Height (m)` * 39.3701, 1))
gym = gym %>%
  rename(`Height (in)` = `Height (m)`)

#converts the gym session duration from hours to minutes
#renames it to "Session_Duration (mins)"
gym = gym %>%
  mutate(`Session_Duration (hours)` = round(`Session_Duration (hours)` * 60, 2))
gym = gym %>%
  rename(`Session_Duration (mins)` = `Session_Duration (hours)`)

#Converts water intake from liters to ounces
#renames it to "Water_Intake (Ounces)"
gym = gym %>%
  mutate(`Water_Intake (liters)` = round(`Water_Intake (liters)` * 33.814, 1))
gym = gym %>%
  rename(`Water_Intake (ounces)` = `Water_Intake (liters)`)

#adds 4 different age groups 
gym = gym %>%
  mutate(Group = case_when(
    Age >= 18 & Age <= 29 ~ "18-29",
    Age >= 30 & Age <= 39 ~ "30-39",
    Age >= 40 & Age <= 49 ~ "40-49",
    Age >= 50 & Age <= 59 ~ "50-59"
  ))

#adds 4 different weight groups 
gym = gym %>%
  mutate(Group_WGT = case_when(
    `Weight (lbs)` >= 88 & `Weight (lbs)` <= 138 ~ "88 - 138",
    `Weight (lbs)` >= 138.01 & `Weight (lbs)` <= 188 ~ "138 - 188",
    `Weight (lbs)` >= 188.01 & `Weight (lbs)` <= 238 ~ "188 - 238",
    `Weight (lbs)` >= 238.01 & `Weight (lbs)` <= 288 ~ "238 - 288"
  ))

```
- Stage 1: EDA 
#creates a boxplot to comapre BMI's of Age Groups
```{r boxplot of age group and BMI}

#code to create an interactive plot for the Age group V. BMI
B1 <- ggplot(data = gym, aes(x = Group, y = BMI, fill = Group)) +
  geom_boxplot() +
  labs(title = "Age Group vs. BMI", x = "Age Group") +
  theme_minimal()

#makes the graph interactive
B1 <- ggplotly(B1)

#plots the graph
B1

```

-  Stage2: Looking to find what is the best formula for a new gym goer to follow:
    -   Weight loss optimization, what works best?
    -   Duration, Workout Type, Water Intake, etc

- What Workout Types are Optimal for Burning Calories?
```{r optimal workouts}

# Get Summary data to display the average of calories burned for workout type
averages <- gym %>%
  group_by(Workout_Type) %>%
  summarize(Average_Cal_Burned = mean(Calories_Burned))

# Code to create lolipop chart that orders the workouts by higher calories burned
lolipop <- ggplot(averages, aes(x = reorder(Workout_Type,-Average_Cal_Burned), y = Average_Cal_Burned)) +
  geom_segment(aes(x = reorder(Workout_Type,-Average_Cal_Burned), 
                   xend = reorder(Workout_Type,-Average_Cal_Burned), 
                   y = 0, yend = Average_Cal_Burned), lwd = 1.5, color = 'Grey') +
    # Creates the lolipop and changes text that pops up when hovering over the lolipops
    # Also changes the color of the lolipop to correlate with workout type
    geom_point(
      aes(color = Workout_Type,
          text = paste(
          "Workout Type: ", Workout_Type,
          "\nAverage Calories Burned: ", round(Average_Cal_Burned, 1)
        )),
        size = 5) +
    labs(
        title = "Average Calories Burned by Workout Type",
        x = "",
        y = "Average Calories Burned",
        ) +
        coord_flip() +
        theme_minimal() +
          theme(legend.position = "none")

# Makes plot interactive
lolipop <- ggplotly(lolipop, tooltip = "text")
lolipop


```

# Scatter Plot for Session Duration and Calories Burned
```{r}

ggplot(gym, aes(x = `Session_Duration (mins)`, y = Calories_Burned, color = as.factor(Experience_Level))) +
  geom_point(alpha = 0.5) +  # Add scatter points
  geom_smooth(method = "loess", se = TRUE, color = "blue") +
  labs(title = "Calories Burned vs. Workout Duration",
       x = "Session Duration (mins)",
       y = "Calories Burned",
       color = "Experience Level") +
  theme_minimal()

```

- Optimal Frequency Based on Experience level, and BMI 
  1 is the lowest, 3 is the highest for Experience level (edit these text chunks later)

```{r}
# Want to look at a lifter who has experience in the gym to see 
# How many days a week they lift and compare to their BMI/Fat%
# Pull in library for sake of comparison
library(patchwork)

# Group gym by experience 
exp_level <- gym %>%
  group_by(Experience_Level) %>%
  select(BMI, `Workout_Frequency (days/week)`)


#Set to a Factor so they will be grouped separately
exp_level$Experience_Level <- as.factor(exp_level$Experience_Level)
exp_level$`Workout_Frequency (days/week)` <- as.factor(exp_level$`Workout_Frequency (days/week)`)


# Creating a violin plot to show distribution of experience level and BMI levels
violin1 <- ggplot(exp_level, aes(x = Experience_Level, y = BMI)) +
  geom_violin(trim = FALSE, aes(fill = Experience_Level),alpha = 0.8) +
  scale_color_gradient(low = "Blue", high = "Red") +
    labs(
      title = "Comparing Experience Level and Workout Frequency",
      x = "Experience Level"
    ) +
  theme(legend.position = "FALSE") 


# Create a stacked bar plot for the sake of comparison
bar1 <- ggplot(exp_level, aes(x = `Workout_Frequency (days/week)`, fill = Experience_Level)) +
  geom_bar(alpha = 0.8, color = "Black") +
  guides(fill = guide_legend(title = "Experience Level")) + 
    labs(
      x = "Workout Frequency (Days per Week)",
      y = "Count"
    ) 


# Using patchwork library to make the graphs be side by side for comparison
comparison <- violin1 + bar1
comparison
```

# Water Intake by Fat %
```{r}
library(plotly)
p = gym %>%
  mutate(Water_Group = cut(`Water_Intake (ounces)`, breaks = c(50, 75, 100, 125, Inf), 
                           labels = c("Low (50 - 75 (oz))", "Medium (75 - 100 (oz))", "High (100 - 125 (oz))", "Very High (125+ (oz))"))) %>%
  ggplot(aes(x = Water_Group, y = Fat_Percentage, fill = Water_Group)) +
  geom_boxplot() +
  labs(title = "Fat Percentage by Water Intake Group",
       x = "Water Intake Group (ounces)",
       y = "Fat Percentage (%)") +
  theme_minimal()

p = ggplotly(p, tooltip = "text")
p

```

# Boxplots for Gender
```{r}

ggplot(gym, aes(x = Gender, y = `Session_Duration (mins)`, fill = Gender)) +
  geom_boxplot(alpha = 0.7) +
  labs(title = "Session Duration by Gender",
       x = "Gender",
       y = "Session Duration (mins)") +
  theme_minimal()


ggplot(gym, aes(x = Gender, y = Calories_Burned, fill = Gender)) +
  geom_boxplot(alpha = 0.7) +
  labs(title = "Calories Burned by Gender",
       x = "Gender",
       y = "Calories Burned") +
  theme_minimal()

```
